### 2023-12-19 ms/tabs/example.tex
updated movement rate .csv files including example .tex to format as latex tables

\documentclass[11pt, oneside]{article}
\usepackage{geometry}
\geometry{letterpaper}

% Load the .csv to table package
\usepackage{pgfplotstable}

\begin{document}

% Example table script
\begin{table}[h]
  \begin{center}
  \caption{Sablefish movement rates between regions (per fish per year).}
  \label{tab:movement-rates}
    \pgfplotstabletypeset[
      col sep=comma,                       % Commas separate .csv values
      columns/{0}/.style={column name={}}, % Replace autofilled column name by " "
      string type                          % Expect characters
    ]{example.csv}
  \end{center}
\end{table}

\end{document}